Scalar_M1

### 2015-05-08 models/Model_Playspace.Rmd
model revs

```{r}
r = seq(from=1, to=5) #Ratings
raw.priors = c(0.003, 0.053, 0.403, 0.387, 0.150) #Measured priors
unif.priors = rep(.2, times=5) #Uniform priors

#Informatitivy data from e5
none.prob = c(1, 0, 0, 0, 0)
some.prob = c(.16, .36, .68, .83, .83)
all.prob = c(.01, .11, .35, .68, .98)

scale.info = c(none.prob, some.prob, all.prob)
scale.matrix = matrix(data = scale.info, nrow = 5, ncol = 3)
scale = c("none", "some", "all")
colnames(scale.matrix) = c(scale)

#Informatitivy data from e5 normalized
n.none.prob = none.prob / sum(none.prob)
n.some.prob = some.prob / sum(some.prob)
n.all.prob = all.prob / sum(all.prob)
n.scale.info = c(n.none.prob, n.some.prob, n.all.prob)
n.scale.matrix = matrix(data = n.scale.info, nrow = 5, ncol = 3)
colnames(n.scale.matrix) = c(scale)
alpha = 4;

#cost function in length of chars
u.cost = function(u) {
  return(nchar(u))
}
```

@@

```{r}

#speaker likelihood helper 
speaker.lhd = function(utt, info) {
  exp(alpha*(log(info) - 3)) #Rething cost??? 
}
#speaker likelihood normalized
speaker.prob = function(utt, rating, s.matrix) {
  num = speaker.lhd(utt, s.matrix[rating, utt])
  norm = 0
  for (i in 1:length(scale) ) {
    current.u = scale[i] #current utterance
    norm = norm + speaker.lhd(current.u, s.matrix[rating, current.u])
  }
  return(num / norm)
}

#non-normalized posterior------------------------------>
nn.post = function(utt, rating, s.matrix, priors) {
  return(speaker.prob(utt, rating, s.matrix) * priors[rating])
}

#normalized posterior---------------------------------->
norm.post = function(utt, rating, s.matrix, priors) {
  nn = nn.post(utt, rating, s.matrix, priors)
  norm = 0
  for (i in 1:5) {
    norm = norm + nn.post(utt, i, s.matrix, priors)
  }
  return(nn / norm)
}

>>>>

Read Data 
---------

```{r}
priors = read.csv("models/prior.csv")
speaker = read.csv("models/speaker.csv")
listener = read.csv("models/listener.csv")
# none.prob = c(1, 0, 0, 0, 0)
# some.prob = c(.16, .36, .68, .83, .83)
# all.prob = c(.01, .11, .35, .68, .98)

@@

```{r}
#cost function in length of chars
u.cost = function(u) {
  return(nchar(u))
}

#speaker likelihood helper 
speaker.lhd = function(utt, info, alpha) {
  exp(alpha*(log(info) - 3)) #Rething cost??? 
}

#speaker likelihood normalized
speaker.prob = function(utt, rating, s.matrix, alpha, scale) {
  num = speaker.lhd(utt, s.matrix[rating, utt], alpha)
  norm = 0
  for (i in 1:length(scale) ) {
    current.u = scale[i] #current utterance
    norm = norm + speaker.lhd(current.u, s.matrix[rating, current.u], alpha)
  }
  return(num / norm)
}

#non-normalized posterior
nn.post = function(utt, rating, s.matrix, priors, alpha, scale) {
  return(speaker.prob(utt, rating, s.matrix, alpha, scale) * priors[rating])
}

#normalized posterior
norm.post = function(utt, rating, s.matrix, priors, 
                     alpha = 1, scale = c("low","hi")) {
  nn = nn.post(utt, rating, s.matrix, priors, alpha, scale)
  norm = 0
  for (i in 1:5) {
    norm = norm + nn.post(utt, i, s.matrix, priors, alpha, scale)
  }
  return(nn / norm)
}

model wrapper functions:

```{r}
run.rsa <- function(d) {

  scale.matrix <- d %>% 
    select(stars, degree, speaker.p, prior.p) %>% 
    spread(degree, speaker.p) %>% 
    mutate(hi = hi / sum(hi), 
           low = low / sum(low)) %>%
    as.matrix

  preds <- expand.grid(stars = 1:5, 
              degree = c("low","hi")) %>%
    rowwise %>%
    mutate(listener.pred =  norm.post(scale, stars, 
                                      s.matrix = scale.matrix[,c("low","hi")], 
                                      prior = scale.matrix[,"prior.p"]))

  return(preds)
} 
```

RUN MODEL
--------

Run model on actual data

```{r}
data <- left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale) %>%
  do(model.preds = run.rsa(.))
